---
title: "431 Class 09"
author: "Thomas E. Love, Ph.D."
date: "2022-09-27"
format:
  revealjs: 
    theme: simple
    self-contained: true
    slide-number: true
    preview-links: auto
    logo: 431-class-foot2.png
    footer: "431 Class 09 | 2022-09-27 | https://thomaselove.github.io/431-2022/"
---

## Today's Agenda

- Pulling in data for a new example, using `read_Rds()`
- Exploring a quantity, broken down into > 2 subgroups
  - Visualization gallery: comparison boxplot, faceted histograms, density and ridgeline plots
- Dealing with missing data via simple (single) imputation
- Using transformations to improve adherence to Normal assumptions, and Tukey's ladder of power transformations

::: aside
Version `r Sys.time()`
:::

## Today's Setup

```{r}
#| echo: true

knitr::opts_chunk$set(comment=NA)
library(broom)               ## tidy up model output
library(equatiomatic)        ## pull equations from regressions
library(ggrepel)             ## build useful labels in ggplot2
library(ggridges)            ## help with ridgeline plots
library(glue)                ## work with strings
library(kableExtra)          ## tidy up tables of output
library(janitor)
library(naniar)
library(simputation)
library(patchwork)
library(tidyverse)

theme_set(theme_bw())
```


## Today's Data

Today, we'll use an R data set (`.Rds`) to import data.

```{r}
#| echo: true
bs_dat <- read_rds("c09/data/blood_storage.Rds")
```

- This allows us to read in the data just as they were last saved in R, including "factoring", etc. 
    - `readRDS()` also works but is a little slower.
- To write an R data set, use `write_rds(datasetname, "locationoncomputer")`. 
    - `saveRDS()` would also work, but slower.

## The blood storage data set {.smaller}

This study^[Cata et al. "Blood Storage Duration and Biochemical Recurrence of Cancer after Radical Prostatectomy". *Mayo Clinic Proceedings* 2011; 86(2): 120-127. <https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3031436/>] evaluates the association between red blood cells (RBC) storage duration (categorized into three groups) and time (in months) to biochemical prostate cancer recurrence after radical prostatectomy. 

In cancer patients, perioperative blood transfusion has long been suspected of reducing long-term survival, and it is suspected that cancer recurrence may be worsened after the transfusion of older blood.

More complete versions of the data (along with more detailed explanations) appear in the [Cleveland Clinic's Statistical Education repository](https://www.lerner.ccf.org/qhs/datasets/), and in the `medicaldata` package in R. 

## Codebook for `bs_dat` (n = 292 subjects)

Variable | Description
------: | :-----------------------------------------------
`participant` | subject identification code
`age_group` | younger, middle or older (RBC age exposure)
`units` | number of allogeneic blood transfusion units received
`recur_time` | time (months) to biochemical recurrence of prostate cancer

Our sample includes participants who received 1-4 units.

## What's in the Data?

```{r}
#| echo: true
bs_dat
```

## Missing Values?

```{r}
#| echo: true
miss_var_summary(bs_dat)
```

## Outcome is time to recurrence

```{r}
#| echo: true
#| warning: false
#| output-location: slide

p1 <- ggplot(bs_dat, aes(sample = recur_time)) +
  geom_qq(col = "dodgerblue") + 
  geom_qq_line(col = "magenta") + 
  theme(aspect.ratio = 1) + 
  labs(title = "Normal Q-Q plot: recur_time")

p2 <- ggplot(bs_dat, aes(x = recur_time)) +
  geom_histogram(aes(y = stat(density)), 
                 bins = 20, fill = "dodgerblue", col = "cyan") +
  stat_function(fun = dnorm, 
                args = list(mean = mean(bs_dat$recur_time, na.rm = TRUE), 
                            sd = sd(bs_dat$recur_time, na.rm = TRUE)),
                col = "magenta", lwd = 1.5) +
  labs(title = "Density Function: recur_time")

p3 <- ggplot(bs_dat, aes(x = recur_time, y = "")) +
  geom_boxplot(fill = "dodgerblue", notch = TRUE, 
               outlier.color = "dodgerblue") + 
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  labs(title = "Boxplot: recur_time", y = "")

p1 + (p2 / p3 + plot_layout(heights = c(4,1)))
```


## Compare `recur_time` by `age_group`

We'll start with a Complete Case Analysis that ignores any case with missing data.

```{r}
#| echo: true
bs_cc <- bs_dat |> filter(complete.cases(age_group, recur_time, units))

mosaic::favstats(recur_time ~ age_group, data = bs_cc) |>
  kbl(digits = 2) |> 
  kable_styling(font_size = 28, full_width = FALSE)
```

## Scatterplot of `recur_time` vs. `age_group`

```{r}
#| echo: true
ggplot(bs_cc, aes(x = age_group, y = recur_time)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
```


## Visualizing Strategies

We're trying to look at the impact of `age_group` on `recur_time`.

- Comparison Boxplot
- Faceted Histograms
- Overlapping Density Plot
- Ridgeline Plot

So let's walk through each of these.

## Comparison Boxplot

```{r}
#| echo: true
#| output-location: slide

ggplot(data = bs_cc, aes(x = age_group, y = recur_time)) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Recurrence Time (in months)",
       x = "Red Blood Cell age group",
       title = "Recurrence Time by RBC Age Group")
```

## Add MEANS to Comparison Boxplot

```{r}
#| echo: true
#| output-location: slide
#| code-line-numbers: 5-6|13|

ggplot(data = bs_cc, aes(x = age_group, y = recur_time)) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Recurrence Time (in months)",
       x = "Red Blood Cell age group",
       title = "Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```

## Faceted Histograms

```{r}
#| echo: true
#| output-location: slide

ggplot(data = bs_cc, aes(x = recur_time, fill = age_group)) +
  geom_histogram(bins = 20, col = "navy") +
  guides(fill = "none") +
  facet_grid(age_group ~ .) +
  labs(x = "Recurrence Time (in months)",
       title = "Recurrence Time by RBC Age Group")
```

## Comparing Densities

```{r}
#| echo: true
ggplot(data = bs_cc, aes(x = recur_time, fill = age_group)) +
  geom_density() + scale_fill_viridis_d(alpha = 0.5, option = "A") + 
  labs(title = "Time to Recurrence, by RBC Age Group")
```

## Using a Ridgeline Plot

```{r}
#| echo: true
ggplot(data = bs_cc, aes(x = recur_time, y = age_group, 
                       fill = age_group)) +
  geom_density_ridges(alpha = 0.5) +
  guides(fill = "none") +
  labs(title = "Time to Recurrence, by RBC Age Group")
```

## Complete Cases: Model Time using Age

```{r}
#| echo: true
m1 <- lm(recur_time ~ age_group, data = bs_cc)

m1
```

## Extract Equation with `equatiomatic`


```{r}
#| echo: true

extract_eq(m1, use_coefs = TRUE, wrap = TRUE, coef_digits = 2,
           terms_per_line = 1, operator_location = "start", 
           font_size = "small")
```

---

```{r}
extract_eq(m1, use_coefs = TRUE, wrap = TRUE, coef_digits = 2,
           terms_per_line = 2, operator_location = "start", 
           font_size = "small")
```


`age_group` | `m1` estimate of `recur_time` (months)
-----------: | -------------:
Younger |
Middle |
Older |

---

```{r}
extract_eq(m1, use_coefs = TRUE, wrap = TRUE, coef_digits = 2,
           terms_per_line = 2, operator_location = "start", 
           font_size = "small")
```


`age_group` | `m1` estimate of `recur_time` (months)
-----------: | -------------:
Younger | 34.29
Middle |
Older |

---

```{r}
extract_eq(m1, use_coefs = TRUE, wrap = TRUE, coef_digits = 2,
           terms_per_line = 2, operator_location = "start", 
           font_size = "small")
```


`age_group` | `m1` estimate of `recur_time` (months)
-----------: | -------------:
Younger | 34.29
Middle | 34.29 - 3.61 = 30.68
Older |

---

```{r}
extract_eq(m1, use_coefs = TRUE, wrap = TRUE, coef_digits = 2,
           terms_per_line = 2, operator_location = "start", 
           font_size = "small")
```


`age_group` | `m1` estimate of `recur_time` (months)
-----------: | -------------:
Younger | 34.29
Middle | 34.29 - 3.61 = 30.68
Older | 34.29 - 0.52 = 33.77

## Sample Means from `bs_cc`

```{r}
#| echo: true

mosaic::favstats(recur_time ~ age_group, data = bs_cc) |>
  select(age_group, mean) |>
  kbl(digits = 2) |> kable_styling(font_size = 28)
```

### Compare to `m1` estimates (some rounding)

`age_group` | Younger | Middle | Older
----: | ----: | ----: | ----:
Est. `recur_time` | 34.29 | 30.68 | 33.77

## Tidy coefficients with `broom` package

```{r}
#| echo: true
tidy(m1, conf.int = TRUE, conf.level = 0.90) |> 
  kbl(digits = 2) |> kable_styling(font_size = 28)
```

- What is the 90% CI for the population mean time to recurrence for `age_group` = Younger?
- What is the 90% CI for the mean difference in time to recurrence between Younger and Middle?

## `glance` to summarize `m1`'s fit

- The `broom` package has three main functions, `tidy()`, `glance()` and `augment()`

```{r}
#| echo: true

glance(m1) |>
  select(r.squared, AIC, nobs, df, df.residual) |>
  kbl(digits = c(4, 1, 0, 0, 0)) |> kable_styling(font_size = 28)
```

# Imputation

## Dealing with the Missing Data

We have done all analyses on complete cases, but that's not always wise.

- What if doing so would bias our conclusions?
- Here we have two missing `age_group` values and one missing `recur_time`.

It's scary to estimate these missing values. What could we do?

## Single Imputation {.smaller}

In single imputation analyses, NA values are estimated/replaced one time with one particular data value for the purpose of obtaining more complete samples, at the expense of creating some potential bias in the eventual conclusions or obtaining slightly less accurate estimates than would be available if there were no missing values in the data.

- The `simputation` package can help us execute single imputations using a wide variety of techniques, within the pipe approach used by the tidyverse. 

See Section 9.8 of the Course Notes for some additional examples.

## Estimate missing values?

```{r}
#| echo: true
bs_dat |> select(-participant) |> summary()
```

Which values are missing and must be imputed?

## Create an imputation model

The `simputation` package is our friend here. We'll use 

- `impute_pmm()` to impute quantities, and 
- `impute_cart()` to impute factors, for now.

```{r}
#| echo: true

bs_imp <- bs_dat |>
  impute_pmm(recur_time ~ age_group + units) |>
  impute_cart(age_group ~ units)
```

We start with no missing `units` so we use that to impute `age_group`, then use both `age_group` and `units` to impute `recur_time`. Any missing data now?

## Compare Results

```{r}
#| echo: true

summary(bs_dat)
summary(bs_imp)
```

## Model Time Using Age with `bs_imp`

```{r}
#| echo: true
m1_imp <- lm(recur_time ~ age_group, data = bs_imp)

extract_eq(m1_imp, use_coefs = TRUE, wrap = TRUE, coef_digits = 2,
           terms_per_line = 1, operator_location = "start", 
           font_size = "small")
```

## Compare Tidied Coefficients

```{r}
#| echo: true
tidy(m1, conf.int = TRUE, conf.level = 0.90) |> 
  kbl(digits = 2) |> kable_styling(font_size = 28)

tidy(m1_imp, conf.int = TRUE, conf.level = 0.90) |> 
  kbl(digits = 2) |> kable_styling(font_size = 28)
```

## Compare Summaries with `glance`

```{r}
#| echo: true
glance(m1) |>
  select(r.squared, AIC, nobs, df, df.residual) |>
  kbl(digits = c(4, 1, 0, 0, 0)) |> kable_styling(font_size = 28)

glance(m1_imp) |>
  select(r.squared, AIC, nobs, df, df.residual) |>
  kbl(digits = c(4, 1, 0, 0, 0)) |> kable_styling(font_size = 28)
```

## What Type of Missingness do we have?

1. MCAR = Missingness completely at random. 

A variable is missing completely at random if the probability of missingness is the same for all units, for example, if for each subject, we decide whether to collect data on a measure by rolling a die and refusing to answer if a “6” shows up. If data are missing completely at random, then throwing out cases with missing data (i.e. doing a complete case analysis) does not bias your inferences.

## What Type of Missingness do we have?

2. MAR = Missingness at random. 

Missingness that depends only on observed predictors. A more general assumption, called missing at random or MAR, is that the probability a variable is missing depends only on available information. Here, we would have to be willing to assume that the probability of nonresponse to depends only on the other, fully recorded variables in the data. 

- Here is the situation that most obviously cries out for imputation.

## What Type of Missingness do we have?

3. Missing not at random

This is a bigger problem, and includes both:

- Missingness that depends on unobserved predictors. Missingness is no longer “at random” if it depends on information that has not been recorded and this information also predicts the missing values. 
- Missingness that depends on the missing value itself. For example, suppose that people with higher earnings are less likely to reveal them.

# OK, back to our Model `m1` with complete cases for the rest of today...

## Save residuals and fitted values for m1

```{r}
#| echo: true

m1_aug <- augment(m1, data = bs_cc)

m1_aug
```

## m1 Residuals vs. Fitted Values

```{r}
#| echo: true

ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() 
```

## Normal Q-Q plot of `m1` Residuals

```{r}
#| echo: true

ggplot(data = m1_aug, aes(sample = .resid)) +
  geom_qq() + geom_qq_line(col = "red")
```

## Back to our Comparison Boxplot

- Does comparing means make sense here? 
- Are these sample distributions "Normal-ish"?

```{r}
ggplot(data = bs_cc, aes(x = age_group, y = recur_time)) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Recurrence Time (in months)",
       x = "Red Blood Cell age group",
       title = "Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```

## Would a Transformation Help Us?

```{r}
#| echo: true
mosaic::favstats(~ recur_time, data = bs_cc)
```

Since all `recur_time` values are positive, we might look at:

$log(time)$, or $1/time$, or $\sqrt{time}$, or $time^2$, for example...

What are we hoping these transformations will do?

## Boxplot 0: `recur_time` by `age_group`

```{r}
ggplot(data = bs_cc, aes(x = age_group, y = recur_time)) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Recurrence Time (in months)",
       x = "Red Blood Cell age group",
       title = "Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```


## Boxplot 1: `log(recur_time)` by `age_group`

```{r}
ggplot(data = bs_cc, aes(x = age_group, y = log(recur_time))) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Natural Log of Recurrence Time",
       x = "Red Blood Cell age group",
       title = "Natural Log of Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```

## Boxplot 2: `1/(recur_time)` by `age_group`

```{r}
ggplot(data = bs_cc, aes(x = age_group, y = 1/(recur_time))) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Inverse of Recurrence Time",
       x = "Red Blood Cell age group",
       title = "Inverse of Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```

## Boxplot 3: $\sqrt{time}$ by `age_group`

```{r}
ggplot(data = bs_cc, aes(x = age_group, y = sqrt(recur_time))) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Square Root of Recurrence Time",
       x = "Red Blood Cell age group",
       title = "Square Root of Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```

## Code for Boxplot 3

```{r}
#| echo: true
#| eval: false

ggplot(data = bs_cc, aes(x = age_group, y = sqrt(recur_time))) +
  geom_violin() +
  geom_boxplot(aes(fill = age_group), width = 0.3, 
               notch = TRUE, outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Square Root of Recurrence Time",
       x = "Red Blood Cell age group",
       title = "Square Root of Recurrence Time by RBC Age Group",
       caption = "Diamonds indicate sample means")
```

## Ridgeline Plot for $\sqrt{time}$?

```{r}
#| echo: true

ggplot(data = bs_cc, aes(x = sqrt(recur_time), y = age_group, 
                       fill = age_group)) +
  geom_density_ridges(alpha = 0.5) +
  guides(fill = "none") +
  labs(title = "Square Root of Time to Recurrence, by RBC Age Group")
```


## Fit a Model to predict $\sqrt{time}$?

```{r}
#| echo: true
m2 <- lm(sqrt(recur_time) ~ age_group, data = bs_cc)

extract_eq(m2, use_coefs = TRUE, wrap = TRUE, coef_digits = 3,
           terms_per_line = 1, operator_location = "start", 
           font_size = "small")
```

## Predicted Values using `m2`

```{r}
extract_eq(m2, use_coefs = TRUE, wrap = TRUE, coef_digits = 3,
           terms_per_line = 1, operator_location = "start", 
           font_size = "small")
```

age_group | Est. $\sqrt{time}$ | Est. `recur_time`
-----: | ----------: | ---------:
Younger | 5.17 | ?
Middle | 5.17 - 0.299 = 4.871 | ?
Older | ? | ?

## Predicted `recur_time` using `m2`

```{r}
extract_eq(m2, use_coefs = TRUE, wrap = TRUE, coef_digits = 3,
           terms_per_line = 1, operator_location = "start", 
           font_size = "small")
```

age_group | Est. $\sqrt{time}$ | Est. `recur_time`
-----: | ----------: | ---------:
Younger | 5.17 | 26.73
Middle | 5.17 - 0.299 = 4.871 | 23.73
Older | 5.17 + 0.014 = 5.184 | 26.87

## Tidy model `m2`

```{r}
#| echo: true

tidy(m2, conf.int = TRUE, conf.level = 0.90) |> 
  kbl(digits = 2) |> kable_styling(font_size = 28)
```

### `glance` to summarize `m2`'s fit

```{r}
#| echo: true

glance(m2) |>
  select(r.squared, AIC, nobs, df, df.residual) |>
  kbl(digits = c(4, 1, 0, 0, 0)) |> kable_styling(font_size = 28)
```

## Normal Q-Q plot of residuals for `m2`

```{r}
#| echo: true
m2_aug <- augment(m2, data = bs_cc)

ggplot(data = m2_aug, aes(sample = .resid)) +
  geom_qq() + geom_qq_line(col = "red")
```


# Power Transformations

## Tukey's Ladder of Power Transformations

- most useful when the outcome is strictly positive
- most useful when dealing with skew in the outcome

Power | -2 | -1 | 0 | 0.5 | 1 | 2 | 3
------: | ----: | ----: | ----: | ----: | ----: | ----: | ----: 
Transformation | $\frac{1}{y^2}$ | $\frac{1}{y}$ | $log(y)$ | $\sqrt{y}$ | $y$ | $y^2$ | $y^3$

- Right Skew usually requires transformations with powers below 1
- Left Skew usually requires powers greater than 1

## Consider the `n_income` data in `dm1000` 

```{r}
#| echo: true

dm1000 <- read_rds("c09/data/dm_1000.Rds")

mosaic::favstats(~ n_income, data = dm1000) |>
  select(n, missing, min, median, mean, max) |>
  kbl(digits = 2) |>
  kable_styling(full_width = FALSE)
```

## Normal Q-Q plot of `n_income`

```{r}
#| echo: true
dm972 <- dm1000 |> filter(complete.cases(n_income))
ggplot(data = dm972, aes(sample = n_income)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1)
```


## Ladder of `n_income` transformations

```{r}
#| echo: true
#| output-location: slide

p1 <- ggplot(data = dm972, aes(sample = n_income)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Untransformed", subtitle = "Power = 1", x = "", y = "")

p2 <- ggplot(data = dm972, aes(sample = n_income^2)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Square", subtitle = "Power = 2", x = "", y = "")

p3 <- ggplot(data = dm972, aes(sample = n_income^3)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Cube", subtitle = "Power = 3", x = "", y = "")

p4 <- ggplot(data = dm972, aes(sample = sqrt(n_income))) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Square Root", subtitle = "Power = 1/2", x = "", y = "")

p5 <- ggplot(data = dm972, aes(sample = log(n_income))) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Logarithm", subtitle = "Power = 0", x = "", y = "")

p6 <- ggplot(data = dm972, aes(sample = 1/n_income)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Inverse", subtitle = "Power = -1", x = "", y = "")

p6 + p5 + p4 + p1 + p2 + p3
```


## Hemoglobin A1c data in `dm1000`

```{r}
#| echo: true
dm985 <- dm1000 |> filter(complete.cases(a1c))
ggplot(data = dm985, aes(sample = a1c)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1)
```

## Ladder of A1c transformations

```{r}
#| echo: true
#| output-location: slide

p1 <- ggplot(data = dm985, aes(sample = a1c)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Untransformed", subtitle = "Power = 1", x = "", y = "")

p2 <- ggplot(data = dm985, aes(sample = a1c^2)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Square", subtitle = "Power = 2", x = "", y = "")

p3 <- ggplot(data = dm985, aes(sample = a1c^3)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Cube", subtitle = "Power = 3", x = "", y = "")

p4 <- ggplot(data = dm985, aes(sample = sqrt(a1c))) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Square Root", subtitle = "Power = 1/2", x = "", y = "")

p5 <- ggplot(data = dm985, aes(sample = log(a1c))) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Logarithm", subtitle = "Power = 0", x = "", y = "")

p6 <- ggplot(data = dm985, aes(sample = 1/a1c)) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Inverse", subtitle = "Power = -1", x = "", y = "")

p7 <- ggplot(data = dm985, aes(sample = 1/(a1c^2))) +
  geom_qq() + geom_qq_line(col = "red") + theme(aspect.ratio = 1) +
  labs(title = "Inverse Square", subtitle = "Power = -2", x = "", y = "")

p7 + p6 + p5 + p4 + p1 + p2
```

# An Example to Work through on your own

## Predict time with `units`

Some data prep required:

- `units` is actually a count.
- Use all 291 observations with `recur_time` and `units`.

```{r}
#| echo: true
bs_dat2 <- bs_dat |>
  filter(complete.cases(recur_time, units))

bs_dat2 |> tabyl(units)
```

## Scatterplot of `recur_time` vs. `age_group`

```{r}
#| echo: true
ggplot(bs_dat2, aes(x = age_group, y = recur_time)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
```


## Comparison Boxplot

```{r}
#| echo: true
#| output-location: slide

ggplot(data = bs_dat2, aes(x = factor(units), y = recur_time)) +
  geom_violin() +
  geom_boxplot(aes(fill = factor(units)), width = 0.3, 
               outlier.size = 2) +
  stat_summary(fun = "mean", geom = "point", 
               shape = 23, size = 3, fill = "white") +
  guides(fill = "none") +
  coord_flip() +
  scale_fill_viridis_d(alpha = 0.5) +
  labs(y = "Recurrence Time (in months)",
       x = "Number of Units Received",
       title = "Recurrence Time by Units",
       caption = "Diamonds indicate sample means")
```

## Model Time using Units

```{r}
#| echo: true

m3 <- lm(recur_time ~ units, data = bs_dat2)

extract_eq(m3, use_coefs = TRUE, coef_digits = 2)

tidy(m3, conf.int = TRUE, conf.level = 0.90)
```

## Model Square Root of Time using Units

```{r}
#| echo: true

m4 <- lm(sqrt(recur_time) ~ units, data = bs_dat2)

extract_eq(m4, use_coefs = TRUE, coef_digits = 2)

tidy(m4, conf.int = TRUE, conf.level = 0.90)
```

## Normal Q-Q plots of Residuals

```{r}
#| echo: true
#| output-location: slide

m3_aug <- augment(m3, data = bs_dat2)
m4_aug <- augment(m4, data = bs_dat2)

p1 <- ggplot(m3_aug, aes(sample = .resid)) +
  geom_qq() + geom_qq_line(col = "red") +
  theme(aspect.ratio = 1) +
  labs(title = "Model m3 Residuals", x = "", y = "")

p2 <- ggplot(m4_aug, aes(sample = .resid)) +
  geom_qq() + geom_qq_line(col = "red") +
  theme(aspect.ratio = 1) +
  labs(title = "Model m4 Residuals", x = "", y = "")

p1 + p2
```

## Compare fits of `m1` and `m3`?

```{r}
#| echo: true

glance(m1) |> select(r.squared, AIC, df, df.residual, nobs)
glance(m3) |> select(r.squared, AIC, df, df.residual, nobs)
```

Are these two models actually predicting the same outcome?

- for the same subjects?


## Session Information

```{r}
#| echo: true
sessionInfo()
```