---
title: "431 Class 17"
author: "Thomas E. Love, Ph.D."
date: "2022-11-01"
format:
  revealjs: 
    theme: simple
    self-contained: true
    slide-number: true
    preview-links: auto
    logo: 431-class-foot2.png
    footer: "431 Class 17 | 2022-11-01 | https://thomaselove.github.io/431-2022/"
---

## Today's Agenda

::: aside
Version `r Sys.time()`
:::

## Today's Packages

```{r}
#| echo: true
#| message: false

library(tidyverse)

theme_set(theme_bw())
```

## Designing a New TB Study

> (PI): OK. That's a nice pilot. We saw $p_{nonshare}$ = 0.18 and $p_{share}$ = 0.26 after your augmentation. Help me design a new study using a two-sided test with $\alpha = 0.05$.

- This time, let's have as many needle-sharers as non-sharers. 
- We should have 90% power to detect a difference almost as large as what we saw in the pilot, or larger, so a difference of 6 percentage points. 

What sample size is required to achieve these aims?

## How `power.prop.test` works

We specify 4 of the following 5 elements of the comparison, and R calculates the fifth.

-   The sample size (interpreted as the \# in each group, so half the total sample size)
-   The true probability in group 1
-   The true probability in group 2
-   The significance level ($\alpha$)
-   The power (1 - $\beta$)

Requires you to work with unbalanced designs.

## Using `power.prop.test`

To find the sample size for a two-sample comparison of proportions using a balanced design:

-   we will use a two-sided test, with $\alpha$ = .05, and power = .90,
-   we estimate that non-sharers have probability .18 of positive tests,
-   and we will try to detect a difference between this group and the needle sharers, who we estimate will have a probability of .24

Any guess as to needed sample size?

## Finding the required sample size in R

```{r}
#| echo: true
power.prop.test(p1 = .18, p2  = .24, alternative = "two.sided",
                sig.level = 0.05, power = 0.90)
```

## Sample Size Required

So, we'd need at least 967 non-sharing subjects, and 967 more who share needles to accomplish the aims of the study, or a total of 1934 subjects.

## Another Scenario

Suppose we can get 400 sharing and 400 non-sharing subjects. How much power would we have to detect a difference in the proportion of positive skin test results between the two groups that was identical to the pilot data above or larger, using a *one-sided* test, with $\alpha$ = .10?

```{r}
#| echo: true
power.prop.test(n=400, p1=.18, p2=.26, sig.level = 0.10,
                alternative="one.sided")
```

## Using the `pwr` package to assess sample size for Unbalanced Designs

The `pwr.2p2n.test` function in the `pwr` package can help assess the power of a test to determine a particular effect size using an unbalanced design, where $n_1$ is not equal to $n_2$.

As before, we specify four of the following five elements of the comparison, and R calculates the fifth.

## Now the five elements are...

-   `n1` = The sample size in group 1
-   `n2` = The sample size in group 2
-   `sig.level` = The significance level ($\alpha$)
-   `power` = The power (1 - $\beta$)
-   `h` = the effect size h, which can be calculated separately in R based on the two proportions being compared: $p_1$ and $p_2$.

## Calculating the Effect Size `h`

To calculate the effect size for a given set of proportions, use `ES.h(p1, p2)` which is available in the `pwr` package.

For instance, comparing .18 to .25, we have the following effect size.

```{r}
#| echo: true
ES.h(p1 = .18, p2 = .25)
```

## Using `pwr.2p2n.test` in R

Suppose we can have 700 samples in group 1 (the not sharing group) but only 400 in group 2 (the group of users who share needles).

How much power would we have to detect the distinction between p1 = .18, p2 = .25 with a 5% significance level in a two-sided test?

## R Command to find the resulting power

```{r}
#| echo: true
pwr::pwr.2p2n.test(h = ES.h(p1 = .18, p2 = .25), n1 = 700, n2 = 400, 
                   sig.level = 0.05)
```

## Comparison to Balanced Design

How does this compare to the results with a balanced design using 1100 drug users in total, i.e. with 550 patients in each group?

```{r}
#| echo: true
pwr::pwr.2p2n.test(h = ES.h(p1 = .18, p2 = .25), n1 = 550, n2 = 550, 
                   sig.level = 0.05)
```

## We could instead have used...

```{r}
#| echo: true
power.prop.test(p1 = .18, p2 = .25, sig.level = 0.05, n = 550)
```


## Not the Same?

Each approach uses approximations, and slightly different ones, so it's not surprising that the answers are similar, but not identical.

### What haven't I included here?

1.  Some people will drop out.
2.  What am I going to do about missing data?
3.  What if I want to do my comparison while adjusting for covariates?

## How Big A Sample Size Do I need?

1.  What is the budget?
2.  What are you trying to compare?
3.  What is the study design?
4.  How big an effect size do you expect (hope) to see?
5.  What was that budget again?
6.  OK, tell me the maximum allowable rates of Type I and Type II error that you want to control for. Or, if you like, tell me the confidence level and power you want to have.
7.  And what sort of statistical inference do you want to plan for?


## Session Information

```{r}
#| echo: true
sessionInfo()
```
